\documentclass{article}
%
\usepackage{amsmath,amssymb,amsthm, tikz-cd}
\usepackage{graphicx}
\usepackage{enumitem}
\usepackage[margin=1in]{geometry}
\usepackage{dutchcal}
\usepackage{colonequals}
\usepackage[
backend=biber,
style=alphabetic,
sorting=ynt
]{biblatex}
\addbibresource{biblio.bib}
%
\usepackage[colorlinks,  citecolor=black, linkcolor = black, urlcolor = black]{hyperref} 
\usepackage{forest}

\usepackage{stmaryrd}
\usepackage{tikz-cd}

%
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}

\theoremstyle{definition}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{recall}[theorem]{Recall}
\newtheorem{example}[theorem]{Example}
\newtheorem{exercise}[theorem]{Exercise}
\newtheorem{solution}{}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem{question}[theorem]{Question}
\newtheorem{problem}[theorem]{Problem}
\newtheorem{notation}[theorem]{Notation}
\newtheorem{setup}[theorem]{Setup}
%
\setlength\parindent{5pt}
\setlength{\parskip}{5pt}
%
\setlist[enumerate]{label = (\roman*), topsep=0pt, itemsep=0pt}

\newlist{enuroman}{enumerate}{1}
\setlist[enuroman]{
	label = (\roman*),
	topsep = 0pt,
	labelindent = \parindent}
	
\newlist{enualph}{enumerate}{2}
\setlist[enualph]{
	label = (\alph*),
	topsep = 0pt,
	labelindent = \parindent}

\newlist{enuarabic}{enumerate}{3}
\setlist[enuarabic]{
	label = (\arabic*),
	topsep = 0pt,
	labelindent = \parindent}

\setlist[itemize]{label = $\diamond$, topsep=0pt, itemsep=0pt, labelindent=0pt}

\newcommand{\comment}[1]{ 
	\begin{center}
		\fcolorbox{red}{white}{
		\begin{minipage}{0.95\textwidth}
			{\color{red} #1}
		\end{minipage}}
	\end{center}}


\newcommand{\C}{\mathbb{C}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\KK}{\mathbb{K}}
\newcommand{\bP}{\mathbb{P}}
\newcommand{\cO}{\mathcal{O}}
\newcommand{\cC}{\mathcal{C}}
\newcommand{\cJ}{\mathcal{J}}
\newcommand{\cM}{\mathcal{M}}
\newcommand{\cN}{\mathcal{N}}
\newcommand{\cF}{\mathcal{F}}
\newcommand{\cG}{\mathcal{G}}
\newcommand{\cH}{\mathcal{H}}
\newcommand{\cA}{\mathcal{A}}
\newcommand{\cB}{\mathcal{B}}
\newcommand{\cQ}{\mathcal{Q}}
\newcommand{\cP}{\mathcal{P}}
\newcommand{\cD}{\mathcal{D}}
\newcommand{\cU}{\mathcal{U}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\la}{\langle}
\newcommand{\ra}{\rangle}
\newcommand{\lp}{\left(}
\newcommand{\rp}{\right)}
\newcommand{\sq}{\subseteq}
\newcommand{\ds}{\dots}
\newcommand{\cds}{\cdots}
\newcommand{\arrow}{arrow}
\newcommand{\F}{\mathbb{F}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\A}{\mathbb{A}}
\newcommand{\fp}{\mathfrak{p}}
\newcommand{\fq}{\mathfrak{q}}
\newcommand{\fm}{\mathfrak{m}}
\newcommand{\fa}{\mathfrak{a}}
\newcommand{\fb}{\mathfrak{b}}
\newcommand{\fc}{\mathfrak{c}}
\newcommand{\fA}{\mathfrak{A}}
\newcommand{\fP}{\mathfrak{P}}
\newcommand{\fn}{\mathfrak{n}}
\newcommand{\bld}[1]{\textbf{#1}}
\newcommand{\raw}{\rightarrow}
\newcommand{\vphi}{\varphi}
\newcommand{\spq}{\supseteq}
\renewcommand{\t}[1]{\widetilde{#1}}
\renewcommand{\b}[1]{\overline{#1}}
\newcommand{\ul}[1]{\underline{#1}}
\renewcommand{\hat}{\widehat}
\newcommand{\ehat}{\hat{ \ }}
\newcommand{\Mustata}{Musta\c{t}\u{a}}
\newcommand{\ceq}{\colonequals}
\newcommand{\coloneqq}{\colonequals}

\DeclareMathOperator{\GL}{GL}
\DeclareMathOperator{\SL}{SL}
\DeclareMathOperator{\Hom}{Hom}
\DeclareMathOperator{\End}{End}
\DeclareMathOperator{\Aut}{Aut}
\DeclareMathOperator{\PGL}{PGL}
\DeclareMathOperator{\Sym}{Sym}
\DeclareMathOperator{\Spec}{Spec}
\DeclareMathOperator{\Frac}{Frac}
\DeclareMathOperator{\Mor}{Mor}
\DeclareMathOperator{\cl}{cl}
\DeclareMathOperator{\id}{id}
\DeclareMathOperator{\Id}{Id}
\DeclareMathOperator{\im}{im}
\DeclareMathOperator{\image}{image}
\DeclareMathOperator{\Ann}{Ann}
\DeclareMathOperator{\Ass}{Ass}
\DeclareMathOperator{\Supp}{Supp}
\DeclareMathOperator{\ini}{in}
\DeclareMathOperator{\Span}{Span}
\DeclareMathOperator{\res}{res}
\DeclareMathOperator{\Proj}{Proj}
\DeclareMathOperator{\rad}{rad}
\DeclareMathOperator{\rank}{rank}
\let\div \relax
\DeclareMathOperator{\div}{div}
\DeclareMathOperator{\val}{val}
\DeclareMathOperator{\gr}{gr}
\DeclareMathOperator{\Syz}{Syz}
\DeclareMathOperator{\Tor}{Tor}
\DeclareMathOperator{\Ext}{Ext}
\DeclareMathOperator{\minpoly}{minpoly}
\DeclareMathOperator{\Fun}{Fun}
\DeclareMathOperator{\FJN}{FJN}
\DeclareMathOperator{\Max}{Max}
\DeclareMathOperator{\coh}{coh}
\DeclareMathOperator{\CD}{CD}
\DeclareMathOperator{\lcm}{lcm}
\DeclareMathOperator{\Irr}{Irr}
\DeclareMathOperator{\Mod}{Mod}
\DeclareMathOperator{\mmod}{-mod}
\DeclareMathOperator{\modm}{mod-}
\DeclareMathOperator{\mumod}{-umod}
\DeclareMathOperator{\mfgumod}{-fgumod}
\DeclareMathOperator{\fpt}{fpt}
\DeclareMathOperator{\Jac}{Jac}
\DeclareMathOperator{\Bl}{Bl}
\DeclareMathOperator{\ac}{ac}
\DeclareMathOperator{\Tr}{Tr}
\DeclareMathOperator{\Char}{char}
\DeclareMathOperator{\Cont}{Cont}
\DeclareMathOperator{\BSR}{BSR}
\DeclareMathOperator{\str}{str}
\DeclareMathOperator{\lct}{lct}
\DeclareMathOperator{\Ten}{Ten}
\DeclareMathOperator{\Gal}{Gal}
\DeclareMathOperator{\Kos}{Kos}
\DeclareMathOperator{\RHom}{RHom}
\DeclareMathOperator{\coker}{coker}
\DeclareMathOperator{\height}{ht}

\renewcommand{\phi}{\varphi}
\renewcommand{\to}{\longrightarrow}
\newcommand{\from}{\longleftarrow}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newcommand{\directory}[1]{
\begin{forest}
  for tree={
    font=\ttfamily,
    grow'=0,
    child anchor=west,
    parent anchor=south,
    anchor=west,
    calign=first,
    edge path={
      \noexpand\path [draw, \forestoption{edge}]
      (!u.south west) +(7.5pt,0) |- node[fill,inner sep=1.25pt] {} (.child anchor)\forestoption{edge label};
    },
    before typesetting nodes={
      if n=1
        {insert before={[,phantom]}}
        {}
    },
    fit=band,
    before computing xy={l=15pt},
  }
#1
\end{forest}
}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%




\newcommand{\code}{\texttt}
	

\begin{document}

\title{Witt vector computations in Macaulay2}
\author{Anne Fayolle, Abhay Goel, Devlin Mallory, Eamon Quinlan-Gallego, Teppei Takamatsu}
\maketitle

\begin{abstract}
We develop an algorithm for performing arithmetic operations in rings of finite length $p$-typical Witt vectors over finitely generated $\F_p$-algebras. We use this to find delta structures on quotients of polynomial rings. We also implement an algorithm to find the quasi-$F$-splitting height of a local complete intersection.
We them implement these algorithm as a package in the computer algebra system Macaulay2. 
\end{abstract}

\comment{
 Funding info.
}


\section{Introduction}

Fix a prime number $p > 0$.

Given a commutative ring $R$ and an integer $n \geq 1$, we let $W_{n}(R)$ be the ring of $p$-typical Witt vectors of length $n$ over $R$. Recall that, as a set, we have $W_n(R) = R \times R \times \cdots \times R$ ($n$ copies), with addition and multiplication defined by certain universal polynomials (see \cite[Section 8.10]{JacobsonBasicAlgebraII} for more details about these universal polynomials). Here are the polynomials for $n=1,2$.
\comment{todo}

The ring structure of $W_n(R)$ is, in fact, engineered so that for every $i = 0, \dots, n-1$ the $i$-th ghost map
%
$$w_i: W_n(R) \to R$$
%
given by 
%
$$w_i(x_0, \dots , x_{n-1}) \ceq x_0^{p^i} + p x_1^{p^{i-1}} + p^2 x_2^{p^{i-2}} + \cdots + p^i x_i$$
%
is a ring homomorphism. 
\comment{Do we need this ghost map in the intro?}


However, these get complicated very fast and make computations quite difficult in $W_n(R)$ for any $n>2$. 
See for instance \cite{FinottiComputationsWittVectorsLength3}. In our implementation, we use a presentation of the ring of Witt vectors as a subalgebra of a polynomial ring as in Illusie \cite{IllusieComplexeDeRhamWittCristalline} (see \autoref{thm-witt-presentation-poly}) to reduce our computations to ones in a polynomial ring. We began our project in Fall 2023 and since then, a similar algorithm has been developped independently by Muñoz-\relax-Bertrand in Sage \cite{MunozBertrandFasterComputationWittVectors}. However, we believe it is still useful to have this algorithm implemented in Macaulay2. 

Using these Witt vector computations, we developped an algorithm that gives constraints to the possible $\delta$-structures on a specific lift $B$ of a given ring $A$ of characteristic $p$ to characteristic $p^2$, see \autoref{section frobeniuslift}. We also implemented an algorithm that allows us to compute the quasi-$F$-splitting height of a complete intersection using the Fedder-type criterion of \cite{KawakamiTakamatsuYoshikawaFedderQuasi}, see \autoref{section quasiFSplitting}.


\comment{Emphasize computations are hard, maybe point to references? Talk about some history? Mention $W(R)$ is the ``universal $\delta$-ring mapping to $R$"? Talk about quasi-$F$-splitting stuff in the abstract?}

\subsection*{Acknowledgements}
We would like to thank Karl Schwede for putting our team together and showing us how to work with Macaulay2. We would also like to thank Kevin Tucker, Jakub Witaszek, and Nawaj KC for useful conversations on this topic

\section{Finite length Witt vectors as finitely generated algebras}

\begin{proposition} \label{prop-witt-presentation-general}
	Let $\tilde R$ be a $p$-torsion free ring for which $R \ceq \tilde R / p \tilde R$ is reduced, and let $n \geq 1$ be an integer. Then the $(n-1)$-th ghost map induces an injective ring homomorphism
	%
	$$\begin{tikzcd}
		w_{n-1}' : W_n(R) \arrow[r, hookrightarrow] & \tilde R / p^n \tilde R
		\end{tikzcd}$$
	%
		whose image is $\tilde R^{p^{n-1}} + p \tilde R^{p^{n-2}} + \cdots + p^{n-1} \tilde R$.
\end{proposition}
%
\begin{proof}
	Since the $(n-1)$-th ghost map $w_{n-1}: W_n(\tilde R) \to \tilde R$ is given by $$w_{n-1}(x_0, \dots , x_{n-1}) = x_0^{p^{n-1}} + p x_1^{p^{n-2}} + p^{n-1} x_0,$$ we see that whenever $x_0, \dots , x_{n-1} \in p \tilde R$ we get $w_{n-1}(x_0, \dots , x_{n-1}) \in p^n \tilde R$. 
	\comment{Not sure how many details to give here: the easiest proof I see is that $p^x > x$ for every real number $x > 0$, and therefore $i + p^{n-1-i} > n$ for every $0 \leq i \leq n-1$} 
	This shows that the composition $W_n(\tilde R) \xrightarrow{w_{n-1}} \tilde R \to \tilde R / p^n \tilde R$ factors through $W_n(R)$ as claimed. The claim about the image is clear from the description of $w_{n-1}$ given above.

	It remains to show that $w_{n-1}'$ is injective; i.e., that given $x_0, \dots, x_n \in \tilde R$ we have $w_{n-1}(x_0, \dots , x_{n-1}) \in p^n \tilde R$ only if $x_0, \dots , x_n \in p \tilde R$. We prove this by induction on $n$, with the base case $n = 1$ being clear.

	Observe that $w_{n-1}(x_0, \dots, x_{n-1}) \equiv x_0^{p^{n-1}} \mod p \tilde R$, and from the assumption that $\tilde R / p \tilde R$ is reduced we obtain that $x_0 \in p \tilde R$, which in turn gives 
	%
	$$0 \equiv w_{n-1}(x_0, \dots , x_{n-1}) \equiv p w_{n-2} (x_1, \dots , x_{n-1}) \mod p^n \tilde R.$$
	%
	Using the $p$-torsion freeness of $\tilde R$, we conclude that $w_{n-2}(x_1, \dots , x_{n-1}) \in p^{n-1} \tilde R$, and now the claim follows from the induction hypothesis.
\end{proof}

\begin{corollary}
	The subset $\tilde R^{p^{n-1}} + p \tilde R^{p^{n-2}} + \cdots + p^{n-1} \tilde R$ is a subring of $\tilde R$.
\end{corollary}

Next, we apply Proposition \ref{prop-witt-presentation-general} to the special case of polynomial rings. This allows us to obtain a presentation of the rings of Witt vectors for such rings, which in turn allows for computations.

\begin{theorem} \label{thm-witt-presentation-poly}
	Let $R \ceq \F_p [X_1, \dots , X_d]$ be a polynomial ring over $\F_p$ and $n \geq 1$ be an integer. Then $W_n(R)$ is isomorphic to the $(\Z / p^n)$-subalgebra of the polynomial ring $(\Z / p^n)[Y_1, \dots , Y_d]$ generated by the elements
	%
	$$ \biggl\{ p^j Y_1^{i_1} Y_2^{i_2} \dots Y_d^{i_d} \ \bigg| \ {{0 \leq j \leq n-1} \atop { 0 \leq i_k \leq p^j}} \biggr\}.$$
\end{theorem}


\begin{proof}
	Applying Proposition \ref{prop-witt-presentation-general} to $\tilde R \ceq \Z [Y_1, \dots , Y_d]$, and identifying $R \cong \tilde R / p \tilde R$ by $X_k = [Y_k \mod p \tilde R]$, we see that it suffices to show that
	\comment{ Finish this proof}

\end{proof}

\begin{remark}
	In the context of Theorem \ref{thm-witt-presentation-poly}, it is useful to think of the variables $Y_k$ as $x_k^{1/p^{n-1}}$.
\end{remark}

\section{Basic examples}


\section{$\delta$-rings and $p$-derivations}\label{section frobeniuslift}
In this section, we explain how we can compute the set of possible $\delta$ structures on a lift $B$ of a given ring $A$ of characteristic $p$ to characteristic $p^2$. 

Giving a $\delta$-structure on $B$ is the same as giving a ring map $\phi\colon A\to W_2(A)$ such that $\pi\circ \phi=\id_A$ where $\pi\colon W_2(A)\to A$ is the projection onto the first coordinate (see \cite[Section 1 - Corollaire]{JoyalDeltaAnneauWitt}).

Based on this, we are able to find equations that the $\delta$-structures on a given ring $B$ must satisfy.

If $A$ is a polynomial ring, say $A=\F_p[x_1,\ldots, x_n]$ then any map $\phi\colon A\to W_2(A)$ as above must send $f(x_1,\ldots, x_n)$ to $(f(x_1,\ldots, x_n),\delta(f(x_1,\ldots, x_n)))$. Note that this implies that $\delta(x_i)$ can be chosen arbitrarily for each i since 
\[
f((x_1,\delta(x_1)),\ldots, (x_n,\delta(x_n)))= (f(x_1,\ldots, x_n),\delta(f(x_1,\ldots, x_n))).
\]
Now, if $A$ is instead a quotient of a polynomial ring, say $A=\F_p[x_1,\ldots,x_n]/(g_1,\ldots, g_r)$, then we again have that any map $\phi\colon A\to W_2(A)$ as above must send $f(x_1,\ldots, x_n)$ to $(f(x_1,\ldots, x_n),\delta(f(x_1,\ldots, x_n)))$ and that 
\[
f((x_1,\delta(x_1)),\ldots, (x_n,\delta(x_n)))= (f(x_1,\ldots, x_n),\delta(f(x_1,\ldots, x_n))).
\]
But in this case, this implies that the $(x_i,\delta(x_i))$ must satisfy the equations defined by the $g_j$s. Note that these are just equations in $W_2(A)$. So if we let $C\coloneqq A[y_1,\ldots,y_n]$ and consider the equations 
\begin{equation}\label{eqn delta is 0}
g_j((x_1,y_1),\ldots, (x_n,y_n))= (g_j(x_1,\ldots, x_n),\delta(g_j(x_1,\ldots, x_n)))=(0,0)
\end{equation}
then the solutions of the $y_j$s in terms of $x_i$s are precisely the possible values of the $\delta(x_i)$s. Here, we abuse notation by seeing $g_j$ as a polynomial in $W_2(C)$ where the $i$-th variable is $(x_i,y_i)$. For instance, if $g_j(x_1,x_2)=x_1^2+x_2$ then $g_j((x_1,y_1),(x_2,y_2))=(x_1,y_1)^2+(x_2,y_2)$. For now, our code lifts element an element $a$ in $\F_p$ to the smallest element $b$ in $\Z/p^2$ with $a=b\mod p$ so $1\in \F_2$ will lift to $1$ in $ \Z/4$, not $3$.


\begin{remark}
If the field over which we are working is not $\F_p$ but rather $\F_q$ for $q$ some power of $p$, we can still pretend we are working in over $\F_p$ by adding variables and adding equations in the defining ideal $(g_1,\ldots,g_r)$ so \autoref{eqn delta is 0} still holds. However, this has to be done by the user, our code only works over $\F_p$. 
\comment{I think we can change this.}
\end{remark}


This is exactly what our \texttt{findFrobeniusLiftConstraints} method does. One can then use it to find examples of Frobenius lifts in different degrees using the \texttt{findFrobeniusLift} method. 

\begin{example} Here is what happens with the cusp for $p=2$.
    \begin{verbatim}
        i1: S=ZZ/2[x,y];
        i2: f=x^2-y^3;
        i3: findFrobeniusLiftConstraints (f)
                     3    2      2     4
        o3 = ideal (y  + x , aa x y + x )
                               1
                       ZZ
        o3 : Ideal of  --[aa ..aa , x..y]
                        2   0    1
        i4 : findFrobeniusLift(f,2)
                        2	
        o4 = {x*y + y, y }
        o4 : List
        i5 : findFrobeniusLift(f,3)
               3    2       2    3    2                 3    2    2
        o5 = {x  + x y + x*y  + y  + x  + x*y + x + y, y  + x  + y }
    \end{verbatim}
\end{example}
This tells us that $aa_0=\delta(x)$ can be anything but $aa_1=\delta(y)$ must be such that $\delta(y)x^2y+x^4=0$. Replacing $x^2$ by $y^3$ gives $\delta(y)y^4+y^6=0$ so $\delta(y)\in y^2 +(f)$. This is exactly what \texttt{findFrobeniusLift(f,2)} returns. Trying for solutions with higher degrees will simply add elements of $(f)$ to the solution as one can see when trying \texttt{findFrobeniusLift(f,3)}.


\section{Quasi $F$-splittings}\label{section quasiFSplitting}
We now explain how to use our code to detect whether or not a given variety is quasi $F$-split and if so, what is its quasi-$F$-splitting height.
We recall the definition of a quasi-$F$-splitting. 
\begin{definition}[\cite{YobukoQuasiFSplittingLiftingCY}]
Let $R$ be a commutative Noetherian ring of characteristic $p>0$ and let $W_n(R)$ be its $n$-th Witt ring for any $n\in \N$. Let $\pi\colon W_n(R)\to R$ be the projection map onto the first coordinate and let $F\colon W_n(R)\to W_n(R)$ be the Frobenius map. The \emph{quasi-$F$-split} height $\height(R)$ of $R$ is the infimum number $n>0$ such that there exists a $W_n(R)$-module homomorphism $\phi$ which makes the following diagram commutative
\begin{center}
\begin{tikzcd}
W_n(R) \arrow[d,"\pi"'] \arrow[r, "F"] & F_* W_n(R) \arrow[dl,dashrightarrow, "\phi"]\\
R
\end{tikzcd}
\end{center}
\end{definition}


In \cite{KawakamiTakamatsuYoshikawaFedderQuasi}, Kawakami, Takamatsu, and Yoshikawa showed the existence of a Fedder-type criterion for quasi-$F$-splittings. We implement a version of it for complete intersections in our package. A faster implementation can be done in Julia, see \cite{BatubaraGarzellaPanFastAlgorithmQuasiFSplitting}. 

\begin{remark}
    We note that the only place our code uses computations in rings of Witt vectors is in the computation of $\Delta_1$ that appears in the theorem \cite[Theorem 5.8]{KawakamiTakamatsuYoshikawaFedderQuasi}. This computation could be done without using Witt vectors, as explained in \cite[Section 1.1]{KawakamiTakamatsuYoshikawaFedderQuasi}. However, we still decided to include it in our package due to the relevance of Witt vectors in quasi-$F$-splitting.
\end{remark}

%This is based on the following Fedder-type criterion  from \cite{KawakamiTakamatsuYoshikawaFedderQuasi}. 
%\begin{theorem}[\cite{KawakamiTakamatsuYoshikawaFedderQuasi}]
%Let $R\coloneqq \F_p\llbracket x_1,\ldots,x_n\rrbracket$. Let $I\coloneqq(f_1,\ldots,f_m)$ be an ideal of $R$ generated by a regular sequence and $f\coloneqq f_1\cdots f_m$.
%\end{theorem}
Let $R=\F_p[x_1,\ldots,x_n]$ and $I\coloneqq(f_1,\ldots,f_m)$ be an ideal of $R$ generated by a regular sequence. 
The method \texttt{fSplittingHeight} 
takes the ideal $I$ and will return the quasi-$F$-splitting height $\height(R/I)$ of $R/I$ if it is less than $102$ or equal to infinity. It will not return anything otherwise. The user should feel free to change this upper limit in their local version of the code.
The limit of $102$ was chosen by us to make the algorithm stop after a reasonable time and we chose $102$ since Artin-Mazur heights of Calabi-Yau quintic threefolds are bounded above by
$102$, when finite. As shown in \cite[Example 7.17]{KawakamiTakamatsuYoshikawaFedderQuasi}, one can construct a variety with arbitrarily high quasi-$F$-splitting height but the one they provide with height $102$ has dimension $2^{51}-1$, making it hard to check on a computer.  


\begin{remark}
The Fedder-type criterion of \cite[Theorem 5.8]{KawakamiTakamatsuYoshikawaFedderQuasi} works for rings defined over any field $k$ of characteristic $p>0$. However, Macaulay2 only works with the $k$-linear Frobenius whereas the Fedder-type criterion uses an absolute Frobenius. Since we want the two to coincide, we only work over $k=\F_p$.
\end{remark}

\begin{example} [\cite{KawakamiTakamatsuYoshikawaFedderQuasi}, Example 7.4] A Calabi-Yau surface of quasi $F$-splitting height $5$.
\begin{verbatim}
i1 : R=ZZ/3[x,y,z,w];

i2 : I=ideal(x^4 + y^4 + z^4 + w^4 + x^3*z + z^3*w + y*z^2*w + y*z*w^2);

            4    4    3     4      2     3         2    4
o2 = ideal(x  + y  + x z + z  + y*z w + z w + y*z*w  + w )

o2 : Ideal of R

i3 : fSplittingHeight(I)

o3 = 5
\end{verbatim}

\end{example}

Let $R\coloneqq \F_p[x_1,\ldots,x_n]$ and $f_1,\ldots,f_r$ be a homogeneous regular sequence in $R$. Let $f\coloneqq f_1\cdots f_n$ and $I\coloneqq(f_1,\ldots,f_n)$. In \cite[Theorem 5.8]{KawakamiTakamatsuYoshikawaFedderQuasi}, they show that if $f^{p-1} \Delta_1(f)^{1+p+\ldots+p^{n-2}}\in \mathfrak{m}^{[p^n]}$ for all $n$, where $\mathfrak{m}$ is the homogeneous maximal ideal of $R$, then the quasi-$F$-splitting height of $R/I$ is infinite. Where $\Delta_1(f)$ is defined such that 
\[
(0,\Delta_1(f))=(f,0)-\Sigma(b_i M_i,0)
\]
in $W_2(S)$ for $f=\Sigma b_i M_i$ with $b_i\in \F_p$ $M_i$ monomials.
By the properties of $\Delta_1(f)$, this is the case if $f^{p-2}\in \mathfrak{m}^{[p]}$ or 
$((f^{p-2})+I^{[p]})f^{p(p-2)}\Delta_1(f) \subset \mathfrak{m}^{[p^2]}$, which is what our code checks. 
We note that there could be a variety with infinite quasi-$F$-splitting height that our code does not check. 

\begin{example} 
\begin{verbatim}
i1 : R=ZZ/3[x,y,z,w];

i2 : I=ideal(x^4 + y^4 + z^4 + w^4);

            4    4    4    4
o2 = ideal(x  + y  + z  + w  )

o2 : Ideal of R

i3 : fSplittingHeight(I)

o3 = infinity
o3 : InfiniteNumber
\end{verbatim}
\end{example}

We do not need the ideal to define a hypersurface. 
\begin{example}
\begin{verbatim}
i1 :  S=ZZ/3[x,y,z,w,u,s];

i2 : I=ideal(x*y + y*w^5 + z*w^3 + x*y*z^3 + y*z*w^2,u*z-z*x^3)

                5        3        2      3           3
o2 = ideal (y*w  + x*y*z  + y*z*w  + z*w  + x*y, - x z + z*u)

              ZZ
o2 : Ideal of --[x..z, w, u, s]
               3


i3 : fSplittingHeight(I)

o3 = 1
\end{verbatim}
\end{example}


\section{Details on the \texttt{Macaulay2} package}

Here is how the different classes should behave at the end of the day, not necessarily that they do so now.

\subsection{The \code{WittOverring} class}

The class \code{WittOverring} is a subclass of \code{PolynomialRing}, with special keys as follows:

\directory{
[WittOverring
  [underring (WittPolynomialRing or WittQuotientRing)]
  [wittSub (RingMap)]
]
}



\subsection{The \code{WittPolynomialRing} class}

Given a \code{PolynomialRing} object, say $R$, and an integer $n \geq 0$, running \code{witt(n, R)} creates an object of the new class \code{WittPolynomialRing}, which is a \code{MutableHashTable} with special entries as follows:


\directory{
[WittPolynomialRing
  [wittLength (ZZ)]
  [unWitt (PolynomialRing)]
  [overring (WittOverring)]
  [explicit (QuotientRing) (!)]
]
}

\code{(!)}: Only computed when necessary.

\subsection{The \code{WittQuotientRing} class}

\directory{
[WittQuotientRing
  [wittLength (ZZ)]
  [ambient (WittPolynomialRing)]
  [unWitt (QuotientRing)]
  [overring (WittOverring)]
  [explicit (QuotientRing) (!)]
]
}

\code{(!)}: Only computed when necessary.


\subsection{The \code{WittRingElement} class}



\code{(!)}: Only computed when necessary.

\subsection{Caching}

The \code{PolynomialRing} and \code{QuotientRing} classes in \code{Macaulay2} have a \code{cache} key, which is used to store data like Grobner bases, etc... 

For a \code{PolynomialRing} or \code{QuotientRing}, say \code{R}, our package (when called appropriately) creates keys \code{R.cache.wittVectors.n} (where $n$ is an integer), which is the \code{WittPolynomialRing} given by \code{witt(n, R)}. 


\printbibliography

%\bibliography{biblio}
%\bibliographystyle{alpha}

\end{document}
